\vspace{1.0cm} \LARGE \bf MERS-CoV spillover at the camel-human interface

### 2017-08-06 mers-structure.tex
Move figures within .tex, add citations, combine figs.

\usepackage{lineno}
\linenumbers

@@

% comments
\usepackage{ulem}
\definecolor{purple}{rgb}{0.459,0.109,0.538}
\def\tb#1#2{\sout{#1} \textcolor{purple}{#2}}
\def\tbc#1{\textcolor{purple}{[#1]}}
\def\gdc#1{\textcolor{blue}{[#1]}}
\def\lmc#1{\textcolor{green}{[#1]}}

@@

Although we find evidence of seasonality in zoonotic spillover timing, no such relationship exists for sizes of human sequence clusters (figure \ref{seasonality}C? \gdc{note - lettering might change}).
This is entirely expected, since little seasonality in human behaviour that could facilitate MERS-CoV transmission is expected following an introduction.
Similarly, we do not observe any trend in human sequence cluster sizes over time (Figure \ref{seasonality}B, Spearman $\rho = 0.06$, $p=0.68$), suggesting that MERS-CoV outbreaks in humans are neither growing nor shrinking in size.
This is not surprising either, since MERS-CoV is a camel virus that has to date, experienced little-to-no selective pressure to improve transmissibility between humans.

@@

@@

@@

% \begin{figure}[h]
% \centering
%
% 	\caption{\textbf{Pairwise co-occurence matrix for human clusters.}
% Heatmap shows the posterior probability that a pair of tips from trees of genomic fragments fall within the same clade - tips from fragment 1 are on the x axis, tips from fragment 2 are on the y axis.
% Tips are ordered by their appearance in tree of genome fragment 2 (positions from nucleotide 21000 onwards) reduced to just the human tips and coloured by inferred host (blue for human, orange for camel) on the left.
% Human clusters are largely well-supported as monophyletic and consistent across trees of both genomic fragments.
% 	}
% 	\label{coocurrence}
% \end{figure}
\begin{figure}%
    \centering
    \includegraphics[width=0.95\textwidth]{figures/mers_recombinant_features.png}
    \caption{\textbf{Recombinant features of MERS-CoV phylogenies.}
    A) Heatmap showing the posterior probability that viruses from trees of different genomic fragments belong to the same introduction event.
    Genomic fragment 1 represents nucleotide positions 1 to 21000 and genomic fragment 2 represents nucleotide positions 21001 to 29364.
    The same set of viruses are arrayed on the $x$-axis and on the $y$-axis; the $x$-axis shows identity of these viruses along genomic fragment 1 and the $y$-axis shows identity of these viruses along genomic fragment 2.
    Viruses are ordered by their appearance in tree of genomic fragment 2 reduced to just the human tips and coloured by inferred host (blue for human, orange for camel) on the left.
    Human clusters are largely well-supported as monophyletic and consistent across trees of both genomic fragments.
    B) Phylogenies derived from genomic fragment 1 (left) and genomic fragment 2 (right), reduced to just the human tips.
  	Identical taxa are connected via coloured lines with colour indicating vertical position in the tree on the right.
  	Branches are coloured by inferred ancestral host state (human in blue, camel in orange).
  	Human clusters in blue change phylogenetic positions between the trees together, with little incongruence within clusters.
  	This is evidence for recombinant viruses generated in the camel reservoir entering human populations.
    }
    \label{recombinant_features}
\end{figure}

@@

Despite lack of convergence, neither of the two demographic reconstructions show evidence of fluctuations in the relative genetic diversity ($N_e \tau$) of  MERS-CoV over time (Figure \ref{skygrid}).
However, we do note that estimates of relative genetic diversity are relatively low overall with a mean estimate of $N_e\tau$ of 3.49 years (95\% HPD: 2.71--4.38), and consequently MERS-CoV phylogeny resembles a ladder often seen in human influenza A virus phylogenies \citep{bedford_strength_2011}.
This estimate of $N_e\tau$ can be translated into an estimate of effective prevalence $N_e$ given an estimate of generation time or serial intervals.
Given a 20 day duration of infection in camels \tbc{please add citation}, we expect a 10 day serial interval or $\tau$ of 0.027 years per generation.
Dividing by $\tau$ yields an estimate of $N_e$ of 129 (100--162) individuals.

Here, we compare this estimate of effective prevalence to a rough estimate of census prevalence.
Given extremely high seroprevalence estimates within camels in Saudi Arabia \tbc{please add citation}, we expect camels to usually be infected within their first year of life.
Therefore we can estimate the rough number of camel infections per year as the number of calves produced each year.
We find there are $830\,000$ camels in Saudi Arabia \citep{abdallah_camel_farming_2013} and that female camels in Saudi Arabia have an average fecundity of 45\% \citep{abdallah_camel_farming_2013}.
Thus, we expect $830\,000 \times 0.50 \times 0.45 = 186\,750$ new calves produced yearly and correspondingly $186\,750$ new infections every year, which spread over 10 day serial intervals gives an average of $N = 5042$ infections.

@@

\tbc{Say something like: ``We find that this result is robust to recombination by...'' and reference Figure \ref{skygrid_comparison}.}

@@

Odds ratios of introductions can then be computed for each month $i$ as $\text{OR}_i = \exp(\beta_i)$.
\tbc{Please spell out what parameters are referring to. In English, what is $\alpha_j$ and what is $\beta_i$? I'm having a hard time figuring out what's going on here.}
\tbc{Can you move `\textit{logit}' to the right hand side of the equation (if not too messy)?}

>>>>

%\usepackage{lineno}
%\linenumbers

@@

% comments
%\usepackage{ulem}
\definecolor{purple}{rgb}{0.459,0.109,0.538}
\def\tb#1#2{\sout{#1} \textcolor{purple}{#2}}
\def\tbc#1{\textcolor{purple}{[#1]}}
\def\gdc#1{\textcolor{blue}{[#1]}}
\def\lmc#1{\textcolor{green}{[#1]}}

@@

@@

Although we find evidence of seasonality in zoonotic spillover timing, no such relationship exists for sizes of human sequence clusters (figure \ref{seasonality}B).
This is entirely expected, since little seasonality in human behaviour that could facilitate MERS-CoV transmission is expected following an introduction.
Similarly, we do not observe any trend in human sequence cluster sizes over time (Figure \ref{seasonality}B, Spearman $\rho = 0.06$, $p=0.68$), suggesting that MERS-CoV outbreaks in humans are neither growing nor shrinking in size.
This is not surprising either, since MERS-CoV is a camel virus that has to date, experienced little-to-no selective pressure to improve transmissibility between humans.

@@

\begin{figure}%
    \centering
    \includegraphics[width=0.95\textwidth]{figures/mers_recombinant_features.png}
    \caption{\textbf{Recombinant features of MERS-CoV phylogenies.}
    A) Heatmap showing the posterior probability that viruses from trees of different genomic fragments belong to the same introduction event.
    Genomic fragment 1 represents nucleotide positions 1 to 21000 and genomic fragment 2 represents nucleotide positions 21001 to 29364.
    The same set of viruses are arrayed on the $x$-axis and on the $y$-axis; the $x$-axis shows identity of these viruses along genomic fragment 1 and the $y$-axis shows identity of these viruses along genomic fragment 2.
    Viruses are ordered by their appearance in tree of genomic fragment 2 reduced to just the human tips and coloured by inferred host (blue for human, orange for camel) on the left.
    Human clusters are largely well-supported as monophyletic and consistent across trees of both genomic fragments.
    B) Phylogenies derived from genomic fragment 1 (left) and genomic fragment 2 (right), reduced to just the human tips.
  	Taxa descended from the same introduction in fragment 1 are connected to their counterparts in fragment 2 with combinations of line dotting and colour unique to each introduction clade.
  	Branches are coloured by inferred ancestral host state (human in blue, camel in orange).
  	Human clusters in blue change phylogenetic positions between the trees together, with little incongruence within clusters.
  	This is evidence for recombinant viruses generated in the camel reservoir entering human populations.
    }
    \label{recombinant_features}
\end{figure}

@@

% \begin{figure}[h]
% \centering
%
% 	\caption{\textbf{Pairwise co-occurence matrix for human clusters.}
% Heatmap shows the posterior probability that a pair of tips from trees of genomic fragments fall within the same clade - tips from fragment 1 are on the x axis, tips from fragment 2 are on the y axis.
% Tips are ordered by their appearance in tree of genome fragment 2 (positions from nucleotide 21000 onwards) reduced to just the human tips and coloured by inferred host (blue for human, orange for camel) on the left.
% Human clusters are largely well-supported as monophyletic and consistent across trees of both genomic fragments.
% 	}
% 	\label{coocurrence}
% \end{figure}

@@

Despite lack of convergence, neither of the two demographic reconstructions show evidence of fluctuations in the relative genetic diversity ($N_e \tau$) of  MERS-CoV over time (Figure \ref{skygrid}).
We recover a similar demographic trajectory when estimating $N_{e}\tau$ over time with a skygrid tree prior across the genome split into ten fragments with independent phylogenetic trees to account for confounding effects of recombination (Figures \ref{skygrid}, \ref{skygrid_comparison}).
However, we do note that estimates of relative genetic diversity are relatively low overall with a mean estimate of $N_e\tau$ of 3.49 years (95\% HPD: 2.71--4.38), and consequently MERS-CoV phylogeny resembles a ladder often seen in human influenza A virus phylogenies \citep{bedford_strength_2011}.
This estimate of $N_e\tau$ can be translated into an estimate of effective prevalence $N_e$ given an estimate of generation time or serial intervals.
Given a 20 day duration of infection in camels \citep{adney_replication_2014}, we expect a 10 day serial interval or $\tau$ of 0.027 years per generation.
Dividing by $\tau$ yields an estimate of $N_e$ of 129 (100--162) individuals.

Here, we compare this estimate of effective prevalence to a rough estimate of census prevalence.
Given extremely high seroprevalence estimates within camels in Saudi Arabia \citep{muller_2014,corman_antibodies_2014,chu_2014,reusken_2013,reusken_2014}, we expect camels to usually be infected within their first year of life.
Therefore we can estimate the rough number of camel infections per year as the number of calves produced each year.
We find there are $830\,000$ camels in Saudi Arabia \citep{abdallah_camel_farming_2013} and that female camels in Saudi Arabia have an average fecundity of 45\% \citep{abdallah_camel_farming_2013}.
Thus, we expect $830\,000 \times 0.50 \times 0.45 = 186\,750$ new calves produced yearly and correspondingly $186\,750$ new infections every year, which spread over 10 day serial intervals gives an average of $N = 5042$ infections.

@@

@@

Odds ratios of introductions can then be computed for each month $i$ as $\text{OR}_i = \exp(\beta_i)$.
%\tbc{Please spell out what parameters are referring to. In English, what is $\alpha_j$ and what is $\beta_i$? I'm having a hard time figuring out what's going on here.}
%\tbc{Can you move `\textit{logit}' to the right hand side of the equation (if not too messy)?}
